---
title: "Dimensionality Reduction"
author: "Anh Thu"
date: "10/18/2021"
output: pdf_document
---

#load the library
```{r}
library(readr)
library(dplyr)
library(plyr)
library(lubridate)
library(ggpubr)
library(tidyr)
library(tidyverse)
library(hrbrthemes)
library(viridis)
library(factoextra)
library(xtable)
library(FactoMineR)
library(cluster.datasets)
```

#Import the data set
```{r}
df <- read.csv("users.db.csv")
colnames(df)
```
#Identifying correlations in the variables
```{r}
#Visual inspection of the data normality using Q-Q plots
ggqqplot(df$score, ylab = "Score")
ggqqplot(df$n.matches, ylab = "Number of matches")
ggqqplot(df$n.updates.photo, ylab = "Number of updated photos")
ggqqplot(df$n.photos, ylab = "Number of photos")
```


##Cor.test score & n.matches
```{r}
cor.test(df$score, df$n.matches, method = "pearson")
```

##Histograms of Score and Matches variable
```{r}
par(mfrow=c(1,2))
hist(df$score,main='Score of the profile',xlab='Score')
hist(df$n.matches,main='Total number of matches',xlab='Number of matches')
```

##Scatter plot for score & n.matches
```{r}
ggscatter(df, x = "score", y = "n.matches", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Score of the profile", ylab = "Total matches")
```
## Cor.test gender & n.photos
```{r}
cor_score_keke <- cor.test(df$gender, df$n.photos, method = "pearson")
cor_score_keke
```

## Cor test gender & photo.keke
```{r}
cor.test(df$gender, df$photo.keke, method ='spearman')
```
## Scatter plot for gender & keke
```{r}
ggscatter(df, x = "gender", y = "photo.keke", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "spearman",
          xlab = "Gender", ylab = "Photo keke found")
```
##The correlations between several variables
```{r}
t1 <- round(cor(cbind(df$score,df$n.matches,df$n.updates.photo,df$n.photos)),2)
xtable(t1)
```

#Dimensionality Reduction
```{r}
simple.fit = lm(gender~score, df)
summary(simple.fit)
```
 
```{r}
multi.fit = lm(gender ~ score + n.photos, df)
summary(multi.fit)
```
```{r}
df_active <- df[,c("score", "n.matches", "n.updates.photo", "n.photos","sent.ana","length.prof")]
head(df_active)
```


```{r}
#df_active <- na.omit(df_active)
res.pca <- PCA(df_active, graph= FALSE)
```

```{r}
summary(res.pca)
```
##Scree
```{r}
fviz_eig(res.pca, addlabels = TRUE, ylim = c(0, 50))
```

##The Individuals factor map

```{r}
fviz_pca_ind(res.pca, col.ind = "cos2", # Color by the quality of representation
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE  # Avoid text overlapping
             )
```
```{r}
df <- df %>% mutate(Gender.c =case_when (gender == 0 ~ "Male",
                                          gender == 1 ~ "Female",
                                          gender == 2 ~ "Other"))
```
```{r}
fviz_pca_ind(res.pca,
             geom.ind = "point", # show points only (nbut not "text")
             col.ind = df$Gender.c, # color by groups
             palette = c("#00AFBB", "#E7B800", "#FC4E07"),
             addEllipses = TRUE, # Concentration ellipses
             legend.title = "Groups",
             select.ind = list(contrib=500)
             )
```


##Variables factor map

```{r}
fviz_pca_var(res.pca,
             col.var = "cos2", # Color by qualities of representation
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )
```
##Biplot

```{r}
library(lares)
fviz_pca_biplot(res.pca, repel = TRUE,
                col.var = "#2E9FDF", # Variables color
                col.ind = "#696969" , # Individuals color,
                select.ind = list(contrib=500)
                )
```
##Table


```{r}
# PCA with function prcomp
pca1 = prcomp(df_active, scale. = TRUE)

# sqrt of eigenvalues
pca1$sdev
```

```{r}
# loadings
xtable(pca1$rotation)
```

##MCA

```{r}
df_mca <- df[,c("Gender.c", "photo.keke", "photo.beach", "voyage","laugh")]
df_mca$Gender.c <- as.factor(df_mca$Gender.c)
df_mca$photo.keke <- as.factor(df_mca$photo.keke)
df_mca$photo.beach <- as.factor(df_mca$photo.beach) 
df_mca$voyage <- as.factor(df_mca$voyage)
df_mca$laugh <- as.factor(df_mca$laugh)
head(df_mca)
```
```{r}
res.mca <- MCA(df_mca, graph = FALSE)
res.mca
```

###Eigenvalues / Variances
```{r}
library("factoextra")
eig.val <- get_eigenvalue(res.mca)
head(eig.val)
```


###To visualize the percentages of inertia explained by each MCA dimensions
```{r}
fviz_screeplot(res.mca, addlabels = TRUE, ylim = c(0, 45))
```

###Biplot
```{r}
fviz_mca_biplot(res.mca, repel = TRUE,
                ggtheme = theme_minimal())
```
###Graph of variables
```{r}
var <-get_mca_var(res.mca)
var
```


```{r}
fviz_mca_var(res.mca, choice = "mca.cor", 
            repel = TRUE, # Avoid text overlapping (slow)
            ggtheme = theme_minimal())
```
```{r}
#Coordinates of variable categories
head(round(var$coord, 2), 4)
```
###Variable categories MCA
```{r}
fviz_mca_var(res.mca, 
             repel = TRUE, # Avoid text overlapping (slow)
             ggtheme = theme_minimal())
```

###Quality of representation of variable categories
```{r}
head(var$cos2, 4)
```

```{r}
# Color by cos2 values: quality on the factor map
fviz_mca_var(res.mca, col.var = "cos2",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"), 
             repel = TRUE, # Avoid text overlapping
             ggtheme = theme_minimal())
```
### Individuals MCA
```{r}
fviz_mca_ind(res.mca, col.ind = "cos2", 
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE, # Avoid text overlapping (slow if many points)
             ggtheme = theme_minimal())
```
###Individuals by groups MCA
```{r}
fviz_mca_ind(res.mca, 
             label = "none", # hide individual labels
             habillage = "Gender.c", # color by groups 
             palette = c("#00AFBB", "#E7B800", "#FC4E07"),
             addEllipses = TRUE, ellipse.type = "confidence",
             ggtheme = theme_minimal())
```





### Visualize the cos2 of row categories
```{r}
library("corrplot")
corrplot(var$cos, is.corr=FALSE)
```

```{r}
# Cos2 of variable categories on Dim.1 and Dim.2
fviz_cos2(res.mca, choice = "var", axes = 1:2)
```
```{r}
head(round(var$contrib,2), 4)
```

```{r}
# Contributions of rows to dimension 1
fviz_contrib(res.mca, choice = "var", axes = 1, top = 15)
# Contributions of rows to dimension 2
fviz_contrib(res.mca, choice = "var", axes = 2, top = 15)
```
#K-means

```{r}
# Elbow method
fviz_nbclust(df_active, kmeans, method = "wss") +
    geom_vline(xintercept = 4, linetype = 2)+
  labs(subtitle = "Elbow method")
# Silhouette method
fviz_nbclust(df_active, kmeans, method = "silhouette")+
  labs(subtitle = "Silhouette method")
# Gap statistic
# nboot = 50 to keep the function speedy. 
# recommended value: nboot= 500 for your analysis.
# Use verbose = FALSE to hide computing progression.
set.seed(123)
fviz_nbclust(df_active, kmeans, nstart = 25,  method = "gap_stat", nboot = 50)+
  labs(subtitle = "Gap statistic method")
```


```{r}
km.res <- kmeans(df_active, 2)
```



```{r}
fviz_cluster(km.res, data = df_active,
             palette = c("#2E9FDF", "#E7B800"), 
             geom = "point",
             ellipse.type = "convex", 
             ggtheme = theme_bw()
             )
```


#Hierarchical Clustering


```{r}
res.pca <- PCA(df_active[1:500,], ncp = 3, graph = FALSE)
# 2. HCPC
res.hcpc <- HCPC(res.pca, graph = FALSE)
```


```{r}
fviz_dend(res.hcpc, 
          cex = 0.9,                     # Taille du text
          palette = "jco",               # Palette de couleur ?ggpubr::ggpar
          rect = TRUE, rect_fill = TRUE, # Rectangle autour des groupes
          rect_border = "jco",           # Couleur du rectangle
          labels_track_height = 1, 
          # Augment l'espace pour le texte
)
```


```{r}
library(FactoMineR)
# Compute PCA with ncp = 3
res.pca <- PCA(df_active, ncp = 3, graph = FALSE)
# Compute hierarchical clustering on principal components
res.hcpc <- HCPC(res.pca, graph = FALSE)
```

```{r}
fviz_dend(res.hcpc)
```



```{r}
plot(res.hcpc, choice = "3D.map")
```


```{r}
# Individuals facor map
fviz_cluster(res.hcpc, geom = "point", main = "Factor map")
```

